---
title: "Data processing in R"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
fontsize: 12pt
always_allow_html: yes
bibliography: ACTL90012.bib
csl: taylor-and-francis-chicago-f.csl
---

The following steps are repeated for each state: 

- Extract the earliest possible date of breach occurrence.
- Select breaches within the time periods for which complete and unbiased data are available.
- Examine breaches with negative reporting delay (i.e., occurrence dates later than discovery/notification, or discovery dates later than notification). Correct dates are retrieved from breach notices. If the breach notice does not contain the correct dates or is not available, the dates are changed to NAs.
- Delete observations with no date of breach/organisation name/reported date/unknown number of state residents affected/ineligible breaches (i.e., the number of state residents affected is below the threshold that triggers the notification obligation). 
- Examine and delete duplicate entries regarding the same breach. 
  - Remove exact duplicates (i.e., same reported date, start and end date of breach, the number of state residents affected). 
  - Match organisation name and date of breach to locate original and supplementary notices. The notice with the earliest reported date is the original notice. Then, update the number of state residents affected in the original notice by totalling the number of state residents in all notices, as supplementary notices provide the additional number of state residents found to be affected later on. Lastly, delete supplementary notices. 

```{r, results='hide', message=FALSE, warning=FALSE}
# load packages
library(readr);library(astsa);library(knitr);library(ggplot2);library(TTR);library(SynthETIC);library(data.table);library(ModelMetrics);library(plotly);library(lubridate);library(stringi);library(FluMoDL);library(ChainLadder);library(MASS);library(AER);library(writexl);library(readxl);library(bbmle);library(MuMIn);library(boot);library(pheatmap);library(gplots);library(kableExtra);library(bookdown);library(ggpubr);library(xtable);library(stargazer);library(forcats);library(rgl);library(htmlwidgets);library("RColorBrewer")
libs<- c ("tm", "plyr", "class", "SnowballC", "dplyr", "caret", "corrplot", "beepr","RWeka","e1071","ggplot2", "datasets", "maps","plotly", "wordcloud")
lapply(libs, library, character.only=T)
```

# California

```{r CA}

# Prepare the data for California (2012Q1 - 2021Q4)

ca_mid=read.csv("CA_R.csv",header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", "NaN", ""))

# extract the earliest date of breach

for (i in 2:ncol(ca_mid)) {
  ca_mid[,i] <- as.Date(ca_mid[,i], "%m/%d/%Y")
}

ca_mid$start_date=ca_mid$Date.s..of.Breach

for (i in 1:nrow(ca_mid)) {
  
  if (is.na(ca_mid$start_date[i]) == FALSE) {
    
    dates = c()
    
    for (j in 3:ncol(ca_mid)){
    dates=c(ca_mid[i,j],dates)
    }
    
    dates = na.omit(dates)
    ca_mid$start_date[i] = min(dates)
    
  }
  
}

# Select data, start with 2608

ca_first=data.frame(ca_mid$Organization.Name,ca_mid$Reported.Date,ca_mid$start_date) #first date
colnames(ca_first)<-c("Organization.Name","Reported.Date","Date.of.Breach")

ca_first=ca_first[-which(ca_first$Date.of.Breach<"2012-01-01"),]
ca_first=ca_first[-which(ca_first$Reported.Date>="2021-12-31"),]

## Examine obs with negative reporting delay, record the correct first date of breach

wrong_date=which((ca_first$Reported.Date-ca_first$Date.of.Breach)<0)
ca_first[which((ca_first$Reported.Date-ca_first$Date.of.Breach)<0),]

ca_first$Date.of.Breach[wrong_date[1]]="2020-12-03"
ca_first$Date.of.Breach[wrong_date[2]]="2020-11-01"
ca_first$Date.of.Breach[wrong_date[3]]=NA
ca_first$Date.of.Breach[wrong_date[4]]="2020-10-24"
ca_first$Date.of.Breach[wrong_date[5]]="2020-02-23"
ca_first$Date.of.Breach[wrong_date[6]]="2018-11-01"
ca_first$Date.of.Breach[wrong_date[7]]="2016-08-10"
ca_first$Date.of.Breach[wrong_date[8]]="2016-09-05"
ca_first$Date.of.Breach[wrong_date[9]]="2015-07-30"
ca_first$Date.of.Breach[wrong_date[10]]="2013-09-24"

## Delete observations with no date of breach/organisation name/reported date/unknown number of state residents affected/ineligible breaches (i.e., the number of state residents affected is below the threshold that triggers the notification obligation)

ca_first=ca_first[-which(is.na(ca_first$Date.of.Breach)),]
ca_first=ca_first[-which(is.na(ca_first$Organization.Name)),]
# no breach without reported date 

rownames(ca_first)<-NULL

## Examine and delete duplicate entries regarding the same breach - 56 duplicates, 2198/2254 are not

### Remove exact duplicates

dup=ca_first[duplicated(ca_first),] # 4 exact duplicates
ca_first=ca_first[-which(duplicated(ca_first)),]

### Match organisation name and date of breach

check = data.frame()

dup=ca_first[duplicated(ca_first[,c(1,3)]),]

for (i in 1:nrow(dup)) {

  same_dob_indexes = which(ca_first$Organization.Name == dup[i,]$Organization.Name & ca_first$Date.of.Breach == dup[i,]$Date.of.Breach)

  if (length(same_dob_indexes) > 1){

  temp <- ca_first[same_dob_indexes,]
  check = rbind(check,temp)

  min_don <- min(temp$Reported.Date)

  origin_index <- which(ca_first$Organization.Name == dup[i,]$Organization.Name & ca_first$Date.of.Breach == dup[i,]$Date.of.Breach & ca_first$Reported.Date == min_don)

  if (length(origin_index) > 1){

    # same date of notification as well
    origin_index <- origin_index[1]

  }

  sup_index <- same_dob_indexes[!same_dob_indexes %in% origin_index]

  # delete supplementary notices
  ca_first <- ca_first[-sup_index,]

  }

  }

```

# Indiana

```{r IN}

In=read.csv("IN_R.csv")
colnames(In)<-c("Name.of.Company","Date.of.Notification","Date.of.the.Breach","IN.Affected","Total.Affected")

In=In[-which(In$Name.of.Company=="Name of Company or Organization"),]
In=In[-which(In$Name.of.Company=="Respondent"),]
In=In[-which(In$Name.of.Company=="Matter: Name"),]

In$Date.of.Notification <- as.Date(In$Date.of.Notification, "%m/%d/%Y")
In$Date.of.the.Breach<- as.Date(In$Date.of.the.Breach, "%m/%d/%Y")
In$IN.Affected <- as.numeric(gsub(",","",In$IN.Affected))
In$Total.Affected <- as.numeric(gsub(",","",In$Total.Affected))

# Select data, start with 6040

In=In[-which(In$Date.of.the.Breach<"2014-01-01"),]
In=In[-which(In$Date.of.Notification>="2021-07-01"),]

## Delete obs with negative reporting delay because we don't have details

In=In[-which((In$Date.of.Notification-In$Date.of.the.Breach)<0),]

## Delete observations with no date of breach/organisation name/reported date/unknown number of state residents affected/ineligible breaches (i.e., the number of state residents affected is below the threshold that triggers the notification obligation)

In=In[-which(is.na(In$Date.of.Notification)),]
In=In[-which(is.na(In$Date.of.the.Breach)),]
In=In[-which(is.na(In$IN.Affected)),]

## Examine and delete duplicate entries regarding the same breach - 42 duplicates, 5322/5364 used

### Remove exact duplicates

dup=In[duplicated(In),] # 13 exact duplicates
In=In[-which(duplicated(In)),]

### Match organisation name and date of breach

check = data.frame()

dup=In[duplicated(In[,c(1,3)]),]

for (i in 1:nrow(dup)) {

  same_dob_indexes = which(In$Name.of.Company == dup[i,]$Name.of.Company & In$Date.of.the.Breach == dup[i,]$Date.of.the.Breach)

  if (length(same_dob_indexes) > 1){

  temp <- In[same_dob_indexes,]
  check = rbind(check,temp)

  min_don <- min(temp$Date.of.Notification)

  origin_index <- which(In$Name.of.Company == dup[i,]$Name.of.Company & In$Date.of.the.Breach == dup[i,]$Date.of.the.Breach & In$Date.of.Notification == min_don)

  if (length(origin_index) > 1){

    # same date of notification as well
    origin_index <- origin_index[1]

  } 

  sup_index <- same_dob_indexes[!same_dob_indexes %in% origin_index]

  # update the number of residents of the original notice with the information in the latest supplementary notices - adding the number of additional state residents affected
  In[origin_index,]$IN.Affected <- sum(temp$IN.Affected)

  # delete supplementary notices
  In <- In[-sup_index,]

  }

  }
```

# Montana

```{r MT}
mt=read.csv("MT_R.csv",header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", "NaN", ""))

mt[,2] <- as.Date(mt[,2], "%m/%d/%Y")
mt[,3] <- as.Date(mt[,3], "%m/%d/%Y")
mt[,4] <- as.Date(mt[,4], "%m/%d/%Y")
mt[,5] <- as.numeric(gsub(",", "", mt[,5]))
mt<-mt[-which(is.na(mt$BUSINESS.NAME)),]

# Select data, start with 3175

mt<-mt[-which(mt$START.OF.BREACH<"2015-10-01"),]
mt<-mt[-which(mt$DATE.REPORTED>"2021-12-31"),]

## Check end date minus start date. For entries that remain unchanged, end of breach is incorrect. 

wrong_date=which((mt$END.OF.BREACH-mt$START.OF.BREACH)<0)
mt[which((mt$END.OF.BREACH-mt$START.OF.BREACH)<0),]

mt$START.OF.BREACH[wrong_date[1]]="2020-11-12"
mt$START.OF.BREACH[wrong_date[2]]="2020-07-27"
mt$START.OF.BREACH[wrong_date[7]]="2019-07-29"
mt$START.OF.BREACH[wrong_date[12]]="2018-11-01"
mt$START.OF.BREACH[wrong_date[13]]="2019-09-26"
mt$START.OF.BREACH[wrong_date[14]]="2018-09-23"
mt$START.OF.BREACH[wrong_date[15]]="2018-08-01"
mt$START.OF.BREACH[wrong_date[16]]="2018-08-01"
mt$START.OF.BREACH[wrong_date[18]]="2018-06-12"
mt$START.OF.BREACH[wrong_date[19]]="2018-02-07"
mt$START.OF.BREACH[wrong_date[20]]="2017-07-01"
mt$DATE.REPORTED[wrong_date[20]]="2017-10-23"

mt=mt[-wrong_date[c(4,17)],] # can't verify start of breach, left with 3173

## Check date reported minus start date. 

wrong_date=which((mt$DATE.REPORTED-mt$START.OF.BREACH)<0)
mt[which((mt$DATE.REPORTED-mt$START.OF.BREACH)<0),]

mt$START.OF.BREACH[wrong_date[1]]="2020-11-01"
mt$DATE.REPORTED[wrong_date[2]]="2019-07-12"

## Delete observations with no date of breach/organisation name/reported date/unknown number of state residents affected/ineligible breaches (i.e., the number of state residents affected is below the threshold that triggers the notification obligation)

mt=mt[-which(is.na(mt$START.OF.BREACH)),]
mt<-mt[-which(is.na(mt$DATE.REPORTED)),]
mt<-mt[-which(is.na(mt$MONTANANS.AFFECTED)),] 

## Examine and delete duplicate entries regarding the same breach - 29 duplicates, 3068/3097 used

### Remove exact duplicates

dup=mt[duplicated(mt),] # 7 exact duplicates
mt=mt[-which(duplicated(mt)),]

### Match organisation name and date of breach

check = data.frame()

dup=mt[duplicated(mt[,c(1,2)]),]

for (i in 1:nrow(dup)) {

  same_dob_indexes = which(mt$BUSINESS.NAME == dup[i,]$BUSINESS.NAME & mt$START.OF.BREACH == dup[i,]$START.OF.BREACH)

  if (length(same_dob_indexes) > 1){

  temp <- mt[same_dob_indexes,]
  check = rbind(check,temp)

  min_don <- min(temp$DATE.REPORTED)

  origin_index <- which(mt$BUSINESS.NAME == dup[i,]$BUSINESS.NAME & mt$START.OF.BREACH == dup[i,]$START.OF.BREACH & mt$DATE.REPORTED == min_don)

  if (length(origin_index) > 1){

    # same date of notification as well
    origin_index <- origin_index[1]

  } 

  sup_index <- same_dob_indexes[!same_dob_indexes %in% origin_index]

  # update the number of residents of the original notice with the information in the latest supplementary notices - adding the number of additional state residents affected
  mt[origin_index,]$MONTANANS.AFFECTED <- sum(temp$MONTANANS.AFFECTED)

  # delete supplementary notices
  mt <- mt[-sup_index,]

  }

  }

```

# Maine 

```{r ME}

# Prepare the data 

## Maine 2010-2018

me_2010to2018 <- read.csv("ME_R_2010_2018.csv")
me_2010to2018$Date.of.the.Breach <- as.Date(me_2010to2018$Date.of.the.Breach)
me_2010to2018$Date.of.Notification <- as.Date(me_2010to2018$Date.of.Notification)
me_2010to2018$ME.Affected <- as.numeric(me_2010to2018$ME.Affected)

## Maine 2018-2020

me_late <- read.csv("ME_R_2018_2020.csv")
me_late$X03_01_04_Date.s..Breach.Occurred..start. <- as.Date(me_late$X03_01_04_Date.s..Breach.Occurred..start.,"%m/%d/%Y")
me_late1<- me_late[-which(is.na(me_late$X03_01_04_Date.s..Breach.Occurred..start.)),]

me_late <- read.csv("ME_R_2018_2020.csv")
me_late$X03_01_04_Date.s..Breach.Occurred..start. <- as.Date(me_late$X03_01_04_Date.s..Breach.Occurred..start.,"%B %d %Y")
me_late2<- me_late[-which(is.na(me_late$X03_01_04_Date.s..Breach.Occurred..start.)),]

me_late<-rbind(me_late1,me_late2)
colnames(me_late)<-c("Name.of.Company","Date.of.the.Breach","Date.of.Notification","ME.Affected","Total.Affected")
me_late$Date.of.Notification <- as.Date(me_late$Date.of.Notification,"%m/%d/%Y")
me_late$ME.Affected <- as.numeric(me_late$ME.Affected)
me_late$Total.Affected<- as.numeric(me_late$Total.Affected)

me<-rbind(me_2010to2018,me_late[,1:4]) #2658

# Select data, start with 2462

me=me[-which(me$Date.of.the.Breach<"2013-01-01"),]
me=me[-which(me$Date.of.Notification>="2020-07-01"),]

## Delete obs with Date.of.Notification-Date.of.the.Breach<0 because we don't have details, left with 2446

wrong_date=which((me$Date.of.Notification-me$Date.of.the.Breach)<0)
me$Date.of.Notification[wrong_date[5]]<-"2017-12-04" #Hayden, Narey & Persich, the date wasn't recognised correctly
me=me[-which((me$Date.of.Notification-me$Date.of.the.Breach)<0),]

## Delete observations with no date of breach/organisation name/reported date/unknown number of state residents affected/ineligible breaches (i.e., the number of state residents affected is below the threshold that triggers the notification obligation); no breach with no organisation name

me=me[-which(is.na(me$Date.of.Notification)),]
me=me[-which(is.na(me$Date.of.the.Breach)),]
me=me[-which(is.na(me$ME.Affected)),]

## Examine and delete duplicate entries regarding the same breach - 38 duplicates, 2248/2286 used

### Remove exact duplicates

dup=me[duplicated(me),] # 6 exact duplicates
me=me[-which(duplicated(me)),]

### Match organisation name and date of breach

check = data.frame()

dup=me[duplicated(me[,c(1,2)]),]

for (i in 1:nrow(dup)) {

  same_dob_indexes = which(me$Name.of.Company == dup[i,]$Name.of.Company & me$Date.of.the.Breach == dup[i,]$Date.of.the.Breach)

  if (length(same_dob_indexes) > 1){

  temp <- me[same_dob_indexes,]
  check = rbind(check,temp)

  min_don <- min(temp$Date.of.Notification)

  origin_index <- which(me$Name.of.Company == dup[i,]$Name.of.Company & me$Date.of.the.Breach == dup[i,]$Date.of.the.Breach & me$Date.of.Notification == min_don)

  if (length(origin_index) > 1){

    # same date of notification as well
    origin_index <- origin_index[1]

  } 

  sup_index <- same_dob_indexes[!same_dob_indexes %in% origin_index]

  # update the number of residents of the original notice with the information in the latest supplementary notices - adding the number of additional state residents affected
  me[origin_index,]$ME.Affected <- sum(temp$ME.Affected)

  # delete supplementary notices
  me <- me[-sup_index,]

  }

  }
```

# North Dakota

```{r ND}

nd=read.csv("ND_R.csv",header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", "NaN", ""))

nd[,3] <- as.Date(nd[,3], "%B %d %Y")
nd[,5] <- as.Date(nd[,5], "%B %d %Y")
nd[,4]<-as.Date(nd[,4]," %B %d %Y")
nd$ND.Residents.Affected<-as.numeric(nd$ND.Residents.Affected)

# extract the earliest date of breach

nd$start_date=nd$Date.of.Breach

for (i in 1:ncol(nd)) {
  hi=nd[i,]
  bye=c(hi[1,3],hi[1,4])
  bye=na.omit(bye)
  nd$start_date[i]=min(bye)
}

# create data frames
nd=data.frame(nd$Business.Name,nd$Date.Reported,nd$start_date,nd$ND.Residents.Affected) #first date
colnames(nd)<-c("Organization.Name","Reported.Date","Date.of.Breach","ND.Residents.Affected")

# Select data

nd=nd[-which(nd$Date.of.Breach<"2019-01-01"),]
nd=nd[-which(nd$Reported.Date>="2022-01-01"),]

## Delete observations with no date of breach/organisation name/reported date/unknown number of state residents affected/ineligible breaches (i.e., the number of state residents affected is below the threshold that triggers the notification obligation)

nd=nd[-which(is.na(nd$Date.of.Breach)),]
nd=nd[-which(is.na(nd$ND.Residents.Affected)),]
# no breach without reported date 

## no breach with negative reporting delay

## Examine and delete duplicate entries regarding the same breach - 2 duplicates, 203/205 used

### Remove exact duplicates

dup=nd[duplicated(nd),] # 1 exact duplicates
nd=nd[-which(duplicated(nd)),]

### Match organisation name and date of breach

check = data.frame()

dup=nd[duplicated(nd[,c(1,3)]),]

for (i in 1:nrow(dup)) {

  same_dob_indexes = which(nd$Organization.Name == dup[i,]$Organization.Name & nd$Date.of.Breach == dup[i,]$Date.of.Breach)

  if (length(same_dob_indexes) > 1){

  temp <- nd[same_dob_indexes,]
  check = rbind(check,temp)

  min_don <- min(temp$Reported.Date)

  origin_index <- which(nd$Organization.Name == dup[i,]$Organization.Name & nd$Date.of.Breach == dup[i,]$Date.of.Breach & nd$Reported.Date == min_don)

  if (length(origin_index) > 1){

    # same date of notification as well
    origin_index <- origin_index[1]

  } 

  sup_index <- same_dob_indexes[!same_dob_indexes %in% origin_index]

  # update the number of residents of the original notice with the information in the latest supplementary notices - adding the number of additional state residents affected
  nd[origin_index,]$ND.Residents.Affected <- sum(temp$ND.Residents.Affected)

  # delete supplementary notices
  nd <- nd[-sup_index,]

  }

  }
```

# Washington

```{r WA}
wa=read.csv("WA_R.csv",header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", "NaN", ""))
wa[,1] <- as.Date(wa[,1], "%m/%d/%Y")
wa[,3] <- as.Date(wa[,3], "%m/%d/%Y")
wa$Number.of.Washingtonians.Affected<-as.numeric(wa$Number.of.Washingtonians.Affected) 

# create data frames
wa=data.frame(wa$Organization.Name,wa$Date.Reported,wa$Date.of.Breach,wa$Number.of.Washingtonians.Affected) #first date
colnames(wa)<-c("Organization.Name","Reported.Date","Date.of.Breach","wa.Residents.Affected")

# Select data

wa=wa[-which(wa$Date.of.Breach<"2015-10-01"),]
wa=wa[-which(wa$Reported.Date>="2022-01-01"),]

## Delete observations with no date of breach/organisation name/reported date/unknown number of state residents affected/ineligible breaches (i.e., the number of state residents affected is below the threshold that triggers the notification obligation)

wa=wa[-which(is.na(wa$Date.of.Breach)),] 
wa=wa[-which(is.na(wa$wa.Residents.Affected)),]
wa=wa[-which(wa$wa.Residents.Affected<500),]

# no breach without reported date

## no obs with negative reporting delay

## Examine and delete duplicate entries regarding the same breach - 5 duplicates, 502/507 used

### Remove exact duplicates

dup=wa[duplicated(wa),] # 4 exact duplicates
wa=wa[-which(duplicated(wa)),]

### Match organisation name and date of breach

check = data.frame()

dup=wa[duplicated(wa[,c(1,3)]),]

for (i in 1:nrow(dup)) {

  same_dob_indexes = which(wa$Organization.Name == dup[i,]$Organization.Name & wa$Date.of.Breach == dup[i,]$Date.of.Breach)

  if (length(same_dob_indexes) > 1){

  temp <- wa[same_dob_indexes,]
  check = rbind(check,temp)

  min_don <- min(temp$Reported.Date)

  origin_index <- which(wa$Organization.Name == dup[i,]$Organization.Name & wa$Date.of.Breach == dup[i,]$Date.of.Breach & wa$Reported.Date == min_don)

  if (length(origin_index) > 1){

    # same date of notification as well
    origin_index <- origin_index[1]

  } 

  sup_index <- same_dob_indexes[!same_dob_indexes %in% origin_index]

  # update the number of residents of the original notice with the information in the latest supplementary notices - adding the number of additional state residents affected
  wa[origin_index,]$wa.Residents.Affected <- sum(temp$wa.Residents.Affected)

  # delete supplementary notices
  wa <- wa[-sup_index,]

  }

  }

```

# Delaware

```{r DE}

de=read.csv("DE_R.csv",header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", "NaN", ""))
de[,2] <- as.Date(de[,2], "%m/%d/%y")
de[,3] <- as.Date(de[,3], "%d/%m/%y")
de$Number.of.Potentially.Affected.Delaware.Residents<-as.numeric(de$Number.of.Potentially.Affected.Delaware.Residents)

colnames(de)<-c("Organization.Name","Date.of.Breach","Reported.Date","DE.Residents.Affected")
row.names(de)<-NULL

# Select data
de=de[-which(de$Date.of.Breach<"2018-04-14"),]

## Delete observations with no date of breach/organisation name/reported date/unknown number of state residents affected/ineligible breaches (i.e., the number of state residents affected is below the threshold that triggers the notification obligation)

de=de[-which(is.na(de$Date.of.Breach)),] 
de=de[-which(is.na(de$DE.Residents.Affected)),]
de=de[-which(de$DE.Residents.Affected<500),]

## Examine obs with negative reporting delay, first date of breach 
de$Date.of.Breach[which((de$Reported.Date-de$Date.of.Breach)<0)]<-"2019-08-31"
de$Reported.Date[which((de$Reported.Date-de$Date.of.Breach)<0)]<-"2019-10-22"


## Examine and delete duplicate entries regarding the same breach - no duplicates

```

# Oregon

```{r OR}
or=read.csv("OR_R.csv",header = TRUE, stringsAsFactors = FALSE, na.strings = c("NA", "NaN", ""))
for (i in 2:ncol(or)) {
  or[,i] <- as.Date(or[,i], "%m/%d/%Y")
}

or$start_date=or$Dates.of.Breach

for (i in 1:ncol(or)) {
  hi=or[i,]
  bye=hi[1,3]
  for (j in 4:8){
    bye=c(hi[1,j],bye)
  }
  bye=na.omit(bye)
  or$start_date[i]=min(bye)
}

# create data frames
or=data.frame(or$Organization,or$Reported.Date,or$start_date) #first date
colnames(or)<-c("Organization.Name","Reported.Date","Date.of.Breach")

# Select data 

or=or[-which(or$Date.of.Breach<"2016-01-01"),]
or=or[-which(or$Reported.Date>="2022-01-01"),]

## Delete observations with no date of breach/organisation name/reported date/unknown number of state residents affected/ineligible breaches (i.e., the number of state residents affected is below the threshold that triggers the notification obligation)

or=or[-which(is.na(or$Date.of.Breach)),]

## no obs with negative reporting delay

## Examine and delete duplicate entries regarding the same breach - no duplicates

```

```{r}

## Output "elig_breaches.csv" which contains observations used in the modelling stage

ca_first$state="CA"
ca_first$State.Residents.Affected=500
ca_data <- subset(ca_first, select=c(1,3,2,4,5))
colnames(ca_data)=c("Name.of.Company", "Date.of.Breach", "Reported.Date", "state", "State.Residents.Affected")

In$state="IN"
in_data <- subset(In, select=c(1,3,2,6,4))
colnames(in_data)=c("Name.of.Company", "Date.of.Breach", "Reported.Date", "state", "State.Residents.Affected")

mt$state="MT"
mt_data<-subset(mt, select=c(1,2,4,6,5))
colnames(mt_data)=c("Name.of.Company", "Date.of.Breach", "Reported.Date", "state", "State.Residents.Affected")

me$state="ME"
me_data<-subset(me, select=c(1,2,3,5,4))
colnames(me_data)=c("Name.of.Company", "Date.of.Breach", "Reported.Date", "state", "State.Residents.Affected")

nd$state="ND"
nd_data<-subset(nd, select=c(1,3,2,5,4))
colnames(nd_data)=c("Name.of.Company", "Date.of.Breach", "Reported.Date", "state", "State.Residents.Affected")

wa$state="WA"
wa_data<-subset(wa, select=c(1,3,2,5,4))
colnames(wa_data)=c("Name.of.Company", "Date.of.Breach", "Reported.Date", "state", "State.Residents.Affected")

de$state="DE"
de_data<-subset(de, select=c(1,2,3,5,4))
colnames(de_data)=c("Name.of.Company", "Date.of.Breach", "Reported.Date", "state", "State.Residents.Affected")

or$state="OR"
or$State.Residents.Affected=250
or_data <- subset(or, select=c(1,3,2,4,5))
colnames(or_data)=c("Name.of.Company", "Date.of.Breach", "Reported.Date", "state", "State.Residents.Affected")

all_data=rbind(ca_data,in_data,mt_data,me_data,nd_data,wa_data,de_data,or_data)

write_csv(all_data,"~/Dropbox/R_Studio/Cyber Paper/elig_breaches.csv")
```


